Characterizing subarctic biomes for land surface modeling of pollution and climate risk

### 2021-05-08 VegOzone.tex
Rephrase Section 3.1

\subsection{Ozone data}
\label{subsec:ozone_data}
To determine the significance of the higher \chem{[O_3]} in 2018 in relation to regional averages and trends, we derive a ground-level \chem{[O_3]} climatology for Svanvik (Section~\ref{sec:stats}). From 1986--1996, ozone monitoring has been operated continuously by NILU at Svanhovd, but no ozone measurements have been conducted before the 2018/19 growing season. Due to probable changes in northern hemispheric tropospheric background \chem{[O_3]} and a reduction of summertime peak values related to ozone precursors which fell under air quality regulations in Europe, we cross-calibrate our data with long-term ozone observations at sites in northern Finland and Sweden.

@@

\subsubsection{Ozone}
\label{subsubsec:clim_ozone}
In the following, we will look at derived ozone climatologies from different perspectives.
The multi-annual means of daily mean ozone, for simplicity here without standard deviations, at observation sites in northern Fennoscandia (refer to Table~\ref{tab:ebas_obs} for details) are shown in Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}. All climatologies except for Svanvik show peak values of about $48\,\unit{ppb}$ in late April. The spring peak values at Svanvik are lower because of the much lower altitude compared to the other monitoring sites. The annual average \chem{[O_3]} \chem{\left<[O_3]\right>} declines throughout the growing season until it reaches its minimum in late summer/fall ($24-30\,\unit{ppb}$). The turning points lie approximately in June and November. At Svanvik, \chem{\left<[O_3]\right>} is $6.6\,\unit{ppb}$ lower than on sites at Jergul/Karasjok (NOR), Esrange (SWE), and Pallas (FIN). The large spread in the case of Svanvik and Janiskoski is due to the lower number of observations.

The correlation of \chem{[O_3]} between Esrange, Pallas, and Jergul/Karasjok is fairly high ($r^2\approx 0.8$, Appendix Fig.~\ref{fig:density_distribution}a, b), hence we can put those data series together to derive a general ozone climatology for northern Fennoscandia. The correlation between Svanvik and Esrange and Pallas is rather low, 0.4 and 0.6, respectively. Thus, Svanvik cannot directly be represented by any measurements taken at either Esrange or Pallas. We shall come back to this in Section~\ref{subsec:ozone_reco}.

In Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}, the two dimensional density distributions of \chem{[O_3]} for the combined northern Fennoscandia data (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}a) and Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) are shown. Darker colors indicate higher probability to observe these values. On top of the density distributions, $10\,\unit{days}$ averages of daily mean \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} are displayed together with $1 \sigma$ uncertainties and SE, respectively. Splines have been fitted through the data for easier inter-comparison with reanalysis data.
From the northern Fennoscandia climatology, one can deduce several features. The build-up of the ozone spring peak takes about $7.5\,\unit{months}$ from minimum to maximum. A sharp increase starts one month ahead of the estimated begin of greening $G_\mathrm{begin}$ (see Section~\ref{subsec:do3se_parameters} for details) and is in coincident with rising temperatures and global irradiance. We find day of the year (\unit{doy}) $\left<G_\mathrm{begin}\right> = (150 \pm 14)$ for Svanvik (Supplement Fig~\ref{fig:greening_season_change_Svanvik}), which translates to the end of May. This finding is in line with more extensive studies \citep[][e.g.,]{IJB:Karlsen2007,AFM:Linderholm2006}. The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} is reached in late April ($100-120\,\unit{doy}$). The depletion of ozone from maximum to minimum takes only $4.5\,\unit{months}$. The minimum occurs at $25\,\unit{ppb}$ in the beginning of September ($240-250\,\unit{doy}$).

The density distribution for Svanvik and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} display a similar pattern as the northern Fennoscandia climatology (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b). The spring peak occurs slightly earlier (at the beginning of April). As Svanvik is located at a lower altitude than the other sites, the earlier decline of the ozone peak may point to a stronger influence of vegetation on the removal of ozone. \chem{CO_2} uptake by coniferous trees can occur as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013}. The average \chem{[O_3]} at Svanvik is lower than at the other sites, as already mentioned, which is in league with a decrease in tropospheric ozone with decreasing altitude. In July--September, ozone is occasionally almost completely depleted during nighttime.

>>>>

\subsection{Ozone data}
\label{subsec:ozone_data}
To determine the significance of the higher \chem{[O_3]} in 2018 in relation to regional averages and trends, we derive a ground-level \chem{[O_3]} climatology for Svanvik (Section~\ref{sec:stats}). From 1986--1996, ozone monitoring has been operated continuously by NILU at Svanhovd, but no ozone measurements have been conducted before the 2018/19 growing season.

@@

\subsubsection{Ozone}
\label{subsubsec:clim_ozone}
In the following, we will derived a ozone climatology (\chem{\left<[O_3]\right>}) for Svanvik and for northern Fennoscadia. To this end, we first calculate a pairwise correlation between the different stations. The correlation of \chem{[O_3]} between Esrange, Pallas, and Jergul/Karasjok is fairly high ($r^2\approx 0.8$, Appendix Fig.~\ref{fig:density_distribution}a, b). Therefore, we select these data to derive a generallized ozone climatology for northern Fennoscandia. The correlation between Svanvik with any of the active monitoring sites in Esrange and Pallas is rather low, 0.4 and 0.6, respectively. Thus, conditions at Svanvik cannot directly be represented by measurements taken at either Esrange or Pallas.

In Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}, two dimensional density distributions of \chem{[O_3]} for the combined northern Fennoscandia data (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}a) and Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) are shown. Darker colors indicate higher probability to observe these values. On top of the density distributions, a $10\,\unit{days}$ average of daily mean \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} is displayed together with $1 \sigma$ uncertainties and SE, respectively. Splines have been fitted through the data for easier inter-comparison.

The generalized Fennoscandic \chem{\left<[O_3]\right>} peaks at $48\,\unit{ppb}$ in late April (spring peak). From the northern Fennoscandia climatology, we deduce several features. \chem{\left<[O_3]\right>} declines throughout the growing season until a minimum is reached in late summer/fall ($24-30\,\unit{ppb}$). The turning points lie approximately in June and November. The build-up of the ozone spring peak takes about $7.5\,\unit{months}$ from minimum to maximum. The pronounced increase in \chem{\left<[O_3]\right>} starts on average one month ahead of the estimated begin of greening ($G_\mathrm{begin}$) (see Section~\ref{subsec:do3se_parameters} for details) and coincidents with rising temperatures and global irradiance. For comparison, we found $\left<G_\mathrm{begin}\right> = (150 \pm 14)$ in units of day of the year (\unit{doy}) for Svanvik (Supplement Fig~\ref{fig:greening_season_change_Svanvik}), which translates to the end of May. This finding is in line with more extensive studies of on-site observations and satellite retrievals \citep[][e.g.,]{IJB:Karlsen2007,AFM:Linderholm2006}. The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} is reached in late April ($100-120\,\unit{doy}$). The depletion of ozone from maximum to minimum takes only $4.5\,\unit{months}$. The minimum occurs at $25\,\unit{ppb}$ in the beginning of September ($240-250\,\unit{doy}$).

The density distribution for Svanvik and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} display a similar pattern as the northern Fennoscandia climatology (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b). The spring peak occurs slightly earlier (at the beginning of April) and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $6.6\,\unit{ppb}$ lower than the Fennoscandic climatology. This is in line with a general decrease in tropospheric ozone with decreasing altitude as Svanvik is located at a lower altitude than the other sites. The earlier decline in \chem{\left<[O_3\right>]} may point to an influence of coniferous vegetation on the removal of ozone since\chem{CO_2} uptake by coniferous trees can occur as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013}. In July--September, ozone is occasionally almost completely depleted during nighttime. This may point to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air.

\chem{\left<[O_3]\right>} is likely low-biased at Svanvik due to an increase in northern hemispheric tropospheric background \chem{[O_3]} since the 1990s. To account for this, we derive a bias correction factor by cross-calibrating with the timeseries taken at Esrange and Pallas (Section~\ref{subsec:anomalies}).
